---
title: "EIA Annual Energy Outlook Projections"
author: "Andrew Leach"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
  always_allow_html: yes
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
  
  
---

```{r basics, cache=FALSE,warning=FALSE,message=FALSE}
library(zoo)
library(lubridate)
library(readxl)
library(scales)
library(grid)
library(gridExtra)
library(janitor)
library(ggpubr)
library(cowplot)
library(patchwork)
library(ggthemes)
library(directlabels)
library(pdfetch)
library(gghighlight)
library(viridis)
library(tidyverse)
library(ggrepel)
library(xml2)
library(rvest)

options(dplyr.summarise.inform = FALSE)

data_fetch<-function(key, cat){
  #key<-KEY
  #cat=476336
  ifelse(cat==999999999,
         url <- paste("https://api.eia.gov/category/?api_key=",
                      key, "&out=xml", sep="" ),
         url <- paste("https://api.eia.gov/category/?api_key=",
                      key, "&category_id=", cat, "&out=xml", sep="" )
  )
  
  #http://api.eia.gov/category/?api_key=YOUR_API_KEY_HERE&category_id=476336
  #url <- paste("https://api.eia.gov/category?api_key=",
  #             key, "&category_id=", cat, "&out=xml", sep="" )
  #https://api.eia.gov/category/?api_key=91b4dca0b858df64a2279d82f71af240&category_id=476336&out=xml
  #https://api.eia.gov/category?api_key=91b4dca0b858df64a2279d82f71af240&category_id=476336&out=xml
  
  x <- read_xml(url)
  doc <- XML::xmlParse(file=x)
  
  
  Parent_Category <- tryCatch(XML::xmlToDataFrame(,stringsAsFactors = F,nodes =
                                               XML::getNodeSet(doc, "//category/parent_category_id")),
                              warning=function(w) FALSE, error=function(w) FALSE)
  Sub_Categories <- XML::xmlToDataFrame(,stringsAsFactors = F,nodes =
                                     XML::getNodeSet(doc, "//childcategories/row"))
  Series_IDs <- XML::xmlToDataFrame(nodes =
                                 XML::getNodeSet(doc, "///childseries/row"),stringsAsFactors = F)
  Categories <- list(Parent_Category, Sub_Categories, Series_IDs)
  names(Categories) <- c("Parent_Category", "Sub_Categories", "Series_IDs")
  Categories
}

 get_children<-function(category_id=476336){
   subs<-data_fetch(KEY,cat=category_id)
   sub_cats<-subs$Sub_Categories
   #build list from sub_cats
   cat_store <- list()
   cat_count<-1
   for (cat in sub_cats$category_id) {
     #cat<-sub_cats$category_id[1]
     series<-data_fetch(KEY,cat=cat)
     cat_store[[cat_count]]<-series$Series_IDs
     cat_count<-cat_count+1
   }
   data.frame(do.call(rbind,cat_store))
 }
 #get_children()
 
 get_series<-function(category_id=476336){
   #series,name,f,units,updated
   subs<-data_fetch(KEY,cat=category_id)
   subs$Series_IDs
 }
 #get_series()
 


pd_fix<-function(data,name){
   data<-data.frame(date=index(data), coredata(data))
   data$date<-ymd(data$date)
   data <- setNames(data, c("date",name)) 
 }
 
EIA_to_DF<-function(series_info){
   data<- pdfetch_EIA(series_info$series_id,KEY)
   pd_fix(data,series_info$name)
   }
 


eia_aeo_comp<-function(start_year=2014,end_year=2022,api_series=".SUP_NA_LFL_NA_DCP_NA_USA_MILLBRLPDY.A",
                   label="US Total Crude Oil Production",
                   units="mmbbl/d",
                   history=FALSE,
                   hist_series="PET.MCRFPUS2.A",
                   hist_conversion=1,
                   hist_year=1950,
                   zero_y=TRUE
                   ){ #use oil as the default
  #testing
  #api_series<-".GEN_NA_ELEP_NA_SLR_PHTVL_NA_BLNKWH.A"
  #start_year=2015
  #end_year=2022
  #label="Oil Production"
  #units="mmbbl/d"
  #hist_series<-"PET.MCRFPUS2.A"
  #hist_conversion=1000
  series<-paste("AEO.",seq(start_year,end_year),".REF",seq(start_year,end_year),api_series,sep="")
  labels<-paste(seq(start_year,end_year)," AEO",sep="")
  elements=end_year-start_year+1
  data<-pd_fix(pdfetch_EIA(series,KEY),labels)%>%
    pivot_longer(-date,names_to = "variable")
  plot<-ggplot(data)+
  geom_line(aes(date,value,group=variable,color=variable,size=variable==paste(end_year,"AEO")),lty="31")+
  #geom_point(data=data %>% filter(variable==paste(end_year,"AEO")),aes(date,value,group=variable,color=variable),size=2.25)+
  scale_y_continuous(breaks=pretty_breaks(),expand=c(0,0))+
  scale_x_date(breaks=pretty_breaks(n=10),expand=c(0,0))+
  scale_color_viridis("",discrete = T,option="A",direction = -1,end = .9)+
  scale_size_manual("",values=c(1,1.5))+
  scale_linetype_manual("",values=c("solid"))+
  theme_minimal()+weekly_graphs()+
    theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)))+
  guides(linetype=guide_legend(order = 1,keywidth = unit(1.6,"cm")),
         size="none",
          #shape = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2),
         #linetype = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2),
         #colour = guide_legend(keywidth = unit(1.6,"cm"),override.aes = list(lty = "11")  ,nrow = 2),
         colour = guide_legend(keywidth = unit(1.6,"cm"),nrow = trunc(elements/6)+1,
                               override.aes = list(size=c(rep(1,elements-1),1.5))),
         #fill = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2)
         NULL)
  if(zero_y)
    plot<-plot+expand_limits(y=0)
  
  #get historical data
  if(history)
    {
    hist_data<-pd_fix(pdfetch_EIA(hist_series,KEY),"Historical data")%>%
      pivot_longer(-date,names_to = "variable")%>%
      mutate(value=value/hist_conversion)%>%
      filter(year(date)>=hist_year)
    plot<-plot+
    geom_line(data=hist_data,aes(date,value,lty="Historical Data"),size=1)+
    labs(y=paste(label," (",units,")",sep=""),x="",
         title=paste("Historical",label,"and EIA AEO Reference Case Projections"),
         caption="Source: Data via EIA AEO, graph by Andrew Leach.")
   }
  
  if(!history){
    plot<-plot+
    labs(y=paste(label," (",units,")",sep=""),x="",
         title=paste("EIA Annual Energy Outlook",label,"Forecasts"),
         caption="Source: Data via EIA AEO, graph by Andrew Leach.")
    }
plot
  
  }

```
   
   



## Oil Production, Trade, and Prices {.tabset .tabset-fade}
### US Oil Production
```{r fig.width=14, fig.height=7, oil_prod, cache=TRUE,warning=FALSE}

eia_aeo_comp(history = TRUE,hist_conversion = 1000)
```

### US Oil Imports
```{r fig.width=14, fig.height=7, oil_imp, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".TRAD_NA_LFL_NA_GIM_NA_USA_MILLBRLPDY.A",label = "US Gross Crude Oil Imports",
                          units = "million barrels per day",
                          history = TRUE,
                          hist_series = "PET.MCRIMUS2.A",
                          hist_conversion = 1000,
                          hist_year=1950)
```

### US Oil Exports
```{r fig.width=14, fig.height=7, oil_exp, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".TRAD_NA_LFL_NA_EXP_NA_USA_MILLBRLPDY.A",label = "US Gross Crude Oil Exports",
                          units = "million barrels per day",
                          history = TRUE,
                          hist_series = "PET.MCREXUS2.A",
                          hist_conversion = 1000,
                          hist_year=2010)
```

### WTI Oil Prices
```{r fig.width=14, fig.height=7, oil_price, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2015,api_series =".PRCE_NA_NA_NA_CR_WTI_USA_NDLRPBRL.A",label = "WTI (Cushing) Nominal Spot Price",
                        units = "$/bbl",
                        history = TRUE,
                        hist_series = "PET.RWTC.M",
                        hist_conversion = 1,
                        hist_year=1990)
```


## Natural Gas Production, Trade, and Prices {.tabset .tabset-fade}
### US Natural Gas Production
```{r fig.width=14, fig.height=7, gas_prod, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".SUP_DPR_NA_NA_NG_TOT_USA_TRLCF.A",label = "US Dry Natural Gas Production",
                          units = "TCF",
                          history = TRUE,
                          hist_series = "NG.N9070US2.M",
                          hist_conversion = 10^6/12,
                          hist_year=1950)
```

### US Natural Gas Imports
```{r fig.width=14, fig.height=7, gas_imp, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2020,api_series =".SUP_IMP_NA_NA_NG_NA_NA_TRLCF.A",label = "US Gross Natural Gas Imports",
                          units = "TCF",
                          history = TRUE,
                          hist_series = "NG.N9100US2.M",
                          hist_conversion = 10^6/12,
                          hist_year=1950)
```

### US Natural Gas Exports
```{r fig.width=14, fig.height=7, gas_exp, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".SUP_EXPT_NA_NA_NG_NA_NA_TRLCF.A",label = "US Gross Natural Gas Exports",
                          units = "TCF",
                          history = TRUE,
                          hist_series = "NG.N9130US2.M",
                          hist_conversion = 10^6/12,
                          hist_year=2010)
```

### Henry Hub Natural Gas Prices
```{r fig.width=14, fig.height=7, gas_price, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".PRCE_HHP_NA_NA_NG_NA_USA_NDLRPMBTU.A",label = "Henry Hub Nominal Spot Price",
                          units = "$/MMBTU",
                          history = TRUE,
                          hist_series = "NG.RNGWHHD.M",
                          hist_conversion = 1,
                          hist_year=1990)
```


### Trade Flows
```{r fig.width=14, fig.height=7, gas_trade,verbose = FALSE, cache=TRUE,warning=FALSE}


#AEO Gas trade

 export_set<-c("Exports : Pipeline Exports to Canada",
               "Exports : Pipeline Exports to Mexico",
               "Exports : Liquefied Natural Gas Exports")
 import_set<-c("Imports : Pipeline Imports from Canada",
               "Imports : Pipeline Imports from Mexico",
               "Imports : Liquefied Natural Gas Imports")
 
 
 
#imports by data_series
#http://api.eia.gov/category/?api_key=YOUR_API_KEY_HERE&category_id=476336

import_series<-get_children(476336)
import_series<-filter(import_series,grepl("U.S. Natural Gas Pipeline Imports From",name)|grepl("U.S. Liquefied Natural Gas Imports,",name),!grepl("Price",name),!grepl("Annual",name))
       
#exports by data series
#http://api.eia.gov/category/?api_key=YOUR_API_KEY_HERE&category_id=476803
export_series<-get_children(476802)
export_series<-filter(export_series,grepl("U.S. Natural Gas Pipeline Exports to",name)|grepl("Liquefied U.S. Natural Gas Exports,",name),!grepl("Price",name),!grepl("Annual",name))

gas_trade<-rbind(import_series,export_series)
gas_trade_data<-EIA_to_DF(gas_trade)
#reset to match AEO Names
names(gas_trade_data)<-c("date",                                                  
                         "Imports : Pipeline Imports from Canada",
                         "Imports : Pipeline Imports from Mexico",
                         "Imports : Liquefied Natural Gas Imports",           
                         "Exports : Pipeline Exports to Canada",  
                         "Exports : Pipeline Exports to Mexico",  
                         "Exports : Liquefied Natural Gas Exports")

#gas_trade_data$`Imports : Pipeline Imports`<-gas_trade_data$`Imports : Pipeline Imports from Canada`+gas_trade_data$`Imports : Pipeline Imports from Mexico`
#gas_trade_data$`Exports : Pipeline Exports`<-gas_trade_data$`Exports : Pipeline Exports to Canada`+gas_trade_data$`Exports : Pipeline Exports to Mexico`

#gas_trade_data$`Pipeline Net Imports`<-gas_trade_data$`Imports : Pipeline Imports`- gas_trade_data$`Exports : Pipeline Exports`
#gas_trade_data$`Liquefied Natural Gas Net Imports`<-gas_trade_data$`Imports : Liquefied Natural Gas Imports` - gas_trade_data$`Exports : Liquefied Natural Gas Exports`

  
gas_trade_data<-gas_trade_data %>% pivot_longer(-date,names_to ="series") 

#make it annual
gas_trade_data<-gas_trade_data %>% mutate(year=year(date)) %>% group_by(year,series) %>%
  summarise(value=12*mean(value,na.rm = T)) %>% ungroup %>% #now annual values based on mean non-na month
  mutate(date=ymd(paste(year,12,31,sep = "-")),year=NULL)
#adjust to TCF per year
gas_trade_data$value<-gas_trade_data$value/10^6

history_dates<-tibble(date=unique(gas_trade_data$date))

gas_trade_data<-gas_trade_data %>% filter(series %in% import_set | series %in% export_set) %>%
  mutate(value=ifelse(series %in% import_set,-1*value,value))



#build AEO list

subs<-data_fetch(KEY,cat=3162260)

#AEO_data$test<-"SUP_IMP_LIQ_NA_NG_NA_NA_TRLCF.A"

AEO_data<-subs$Series_IDs
AEO_data<-NULL
for(j in seq(2014,2019)){
  #print(paste("working on AEO",j,sep=""))
  work_data<-subs$Series_IDs
  work_data$series_id<-gsub("2019",j,work_data$series_id)
  work_data$name<-gsub("2019",j,work_data$name)
  AEO_data<-rbind(AEO_data,work_data)
}

#get the 2022 defns
subs<-data_fetch(KEY,cat=4442481)
for(j in seq(2020,2022)){
  #print(paste("working on AEO",j,sep=""))
  work_data<-subs$Series_IDs
  work_data$series_id<-gsub("2022",j,work_data$series_id)
  work_data$name<-gsub("2022",j,work_data$name)
  AEO_data<-rbind(AEO_data,work_data)
}


AEO_data$name<-gsub("Natural Gas : Volumes : ","",AEO_data$name)


test<- pdfetch_EIA(AEO_data$series_id,KEY)
series_data<-data.frame(date=index(test), coredata(test))
series_data$date<-ymd(series_data$date)
series_data <- setNames(series_data, c("date",AEO_data$name))


suppressMessages(series_data <-series_data %>% full_join(history_dates))#include the dates for which we have history. They will be NA now, but we'll stack them in later


#melt it
series_data<-series_data %>% pivot_longer(-c(date),names_to="variable") #%>% na.omit()


#series_data$`Pipeline Net Imports`<-series_data$`Imports : Pipeline Imports`- gas_trade_data$`Exports : Pipeline Exports`
#series_data$`Liquefied Natural Gas Net Imports`<-gas_trade_data$`Imports : Liquefied Natural Gas Imports` - gas_trade_data$`Exports : Liquefied Natural Gas Exports`



#get the year for each
name_split<-do.call(rbind,strsplit(as.character(series_data$variable),", "))

series_data$series<-name_split[,1]
#series_data$case<-name_split[,2]
series_data$aeo_year<-name_split[,3]




suppressMessages(joint_data<-series_data%>%filter(year(date)>=2000)%>%left_join(gas_trade_data%>%rename(history=value)))

joint_data<-joint_data %>% filter(series %in% import_set | series %in% export_set) %>%
  mutate(value=ifelse(series %in% import_set,-1*value,value),
         #history=ifelse(series %in% import_set,-1*history,history)
         NULL)

#strip out history after forecast date

joint_data<-joint_data %>% 
  mutate(year=as.numeric(gsub("AEO","",aeo_year)),
    history=ifelse(year(date)<as.numeric(gsub("AEO","",aeo_year)),history,NA),
    series=gsub("Exports : ","",series),
    series=gsub("Imports : ","",series),
    series=factor(series,levels=c("Liquefied Natural Gas Exports", "Pipeline Exports to Canada",  "Pipeline Exports to Mexico",
                                  "Liquefied Natural Gas Imports", "Pipeline Imports from Canada" ,"Pipeline Imports from Mexico" ))
    
    )


gas_trade_plot<-ggplot(joint_data)+
  geom_area(aes(date,value,fill=series),position="stack",alpha=0.6,color="black",size=0.25)+
  geom_area(aes(date,history,fill=series),position="stack",color="black",size=0.25)+
  facet_wrap(~aeo_year,nrow = 1)+
  scale_y_continuous(breaks=pretty_breaks(),expand=c(0,0))+
  scale_x_date(breaks=pretty_breaks(n=5),expand=c(0,0))+
  #scale_color_viridis("",discrete = T,option="A",direction = -1,end = .9)+
  scale_fill_viridis("",discrete = T,option="A",direction = -1,end = .9)+
  scale_size_manual("",values=c(1,1.5))+
  scale_linetype_manual("",values=c("solid"))+
  expand_limits(x=ymd("1999-10-01"))+
  theme_minimal()+weekly_graphs()+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)))+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust = 0.5))+
  guides(linetype=guide_legend(order = 1,keywidth = unit(1.6,"cm")),
         size="none",
         #shape = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2),
         #linetype = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2),
         #colour = guide_legend(keywidth = unit(1.6,"cm"),override.aes = list(lty = "11")  ,nrow = 2),
         fill = guide_legend(keywidth = unit(1,"cm"),nrow = 2,byrow = TRUE),
         #fill = guide_legend(keywidth = unit(1.6,"cm"),nrow = 2)
         NULL)+
  labs(y=paste("Annual Net Outflows (TCF)",sep=""),x="",
       title=paste("EIA Annual Energy Outlook Natural Gas Trade Projections"),
       subtitle=paste("Historical data up to forecast date in darker shade, forecasts shown with more transparency"),
       caption="Source: Data via EIA, graph by Andrew Leach.")

gas_trade_plot


```




## Electricity supply {.tabset .tabset-fade}


### Total Electricity Supply
```{r fig.width=14, fig.height=7, scheme_plot_3, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2015,api_series =".GEN_NA_ELEP_NA_TEG_NA_USA_BLNKWH.A",label = "Electricity Generation",
                          units = "billion kWh",
                          history = TRUE,
                          hist_conversion = 1000,
                          hist_series = "ELEC.GEN.ALL-US-99.A"
                          )
```


### Solar Generation
```{r fig.width=14, fig.height=7, scheme_plot_1, cache=TRUE,warning=FALSE}

eia_aeo_comp(start_year = 2014,api_series =".GEN_NA_ALLS_NA_SLR_NA_NA_BLNKWH.A",label = "Solar Electricity Generation",
                          units = "billion kWh",
                          history = TRUE,
                          hist_conversion = 1000,
                          hist_series = "ELEC.GEN.TSN-US-99.A"
                          )
```


### Coal Generation

```{r fig.width=14, fig.height=7, scheme_plot_2, cache=TRUE,warning=FALSE}
eia_aeo_comp(start_year = 2018,api_series =".GEN_NA_ELEP_POW_CL_NA_USA_BLNKWH.A",label = "Coal-fired Electricity Generation",
                          units = "billion kWh",
                         history = TRUE,
                         hist_conversion = 1000,
                         hist_series = "ELEC.GEN.COW-US-98.A")
```

## Energy-related carbon dioxide emissions {.tabset .tabset-fade}
```{r fig.width=14, fig.height=7, co2_plot, cache=TRUE,warning=FALSE}
eia_aeo_comp(start_year = 2015,api_series =".EMI_CO2_TEN_NA_NA_NA_NA_MILLMETNCO2.A",label = "Energy-related CO2 Emissions",
                         units = "Mt",
                         history = TRUE,
                         hist_series = "EMISS.CO2-TOTV-TT-TO-US.A",
                        zero_y = TRUE)+expand_limits(y=6000)
```

